---
title: "ESC403_SinusPred"
author: "Jerome Sepin"
date: "14 4 2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Preparation
```{r Preparation}
# Clean Data
rm(list=ls())

library(ggplot2)
library(dplyr)
library(forecast)
```

```{r DataGeneration, echo = FALSE}
# Create Data (deterministic)
days <- 40+10
x <-  c(0:(24*days))
y <-  sin(x/(24/(2*pi) )) + 0.005*x^(1.4) #+ rnorm(length(x), mean = 0, sd = 0.1)
plot(x,y,type = "l",xlab = "Days",ylab = "Temperature...")
length(y)
```


#ARIMA-forecasting
StepPred <- 20
ts_y <- ts(y[1:(length(y)-StepPred)], frequency=24)#need to be known!
decomp <- stl(ts_y, s.window = 'periodic')
plot(decomp)
automaticARIMA <- auto.arima(decomp$time.series[,"remainder"],trace = T, seasonal =F,allowmean=F)#seasonality is off and zeromean here (bc stationary)
futurVal <- forecast(automaticARIMA,h=StepPred,level=c(95))#h=24: 24 values in future
plot(futurVal, xlim = c(45,55))


```{r}
StepPred <- 20
ts_y <- ts(y[1:(length(y)-StepPred)], frequency=24)#need to be known!
fitARIMA=auto.arima(ts_y,trace = F,D=NA) # automatic&forced seasonality (D=NA is default, D=1 is forced for seasonality)
futurVal <- forecast(fitARIMA,h=StepPred,level=c(95))#h=20: 20 values in future

plot(as.numeric(futurVal$mean))s
lines(c(futurVal$lower))
lines(c(futurVal$upper))
lines(y[(length(y)-StepPred+1):length(y)],col= "red")
```


```{r}
true_uprEnd <- y[(length(y)-StepPred+1):length(y)]
pred_uprEnd <- as.numeric(futurVal$mean)

df_prediction <- data.frame(true_uprEnd, pred_uprEnd)
df_prediction$x <- 1:nrow(df_prediction)
df_prediction <- tidyr::gather(df_prediction, key = "key", value = "value",-c(x))
df_prediction <- cbind(df_prediction, matrix(unlist(strsplit(df_prediction$key,"_")),ncol = 2,byrow = T))
df_prediction <- dplyr::select(df_prediction,-c(key))
colnames(df_prediction) <- c("x","y","True","End")


ggplot(data = df_prediction, aes(x=x,y=y,col = as.factor(True) ))+
  geom_line()+
  labs(title = "",y = "",
       x="")+
  facet_wrap(~End,scales = "free")+
  theme_bw()+
  scale_color_manual("",values=c("darkred","darkblue"))+
  theme(legend.position="top")
```





```{r}
climate <- read.csv("../data/pollution.csv")
y <- climate$TEMP
plot(y,type = "l")
```

```{r}
StepPred <- 20
ts_y <- ts(y[1:(length(y)-StepPred)], frequency=24)#need to be known!
fitARIMA=auto.arima(ts_y,trace = F,D=NA) # automatic&forced seasonality (D=NA is default, D=1 is forced for seasonality)
futurVal <- forecast(fitARIMA,h=StepPred,level=c(95))#h=20: 20 values in future

plot(as.numeric(futurVal$mean))
lines(c(futurVal$lower))
lines(c(futurVal$upper))
lines(y[(length(y)-StepPred+1):length(y)],col= "red")
```


```{r}
true_uprEnd <- y[(length(y)-StepPred+1):length(y)]
pred_uprEnd <- as.numeric(futurVal$mean)

MSE <- sum((true_uprEnd-pred_uprEnd)^2)

df_prediction <- data.frame(true_uprEnd, pred_uprEnd)
df_prediction$x <- 1:nrow(df_prediction)
df_prediction <- tidyr::gather(df_prediction, key = "key", value = "value",-c(x))
df_prediction <- cbind(df_prediction, matrix(unlist(strsplit(df_prediction$key,"_")),ncol = 2,byrow = T))
df_prediction <- dplyr::select(df_prediction,-c(key))
colnames(df_prediction) <- c("x","y","True","End")



ggplot(data = df_prediction, aes(x=x,y=y,col = as.factor(True) ))+
  geom_line()+
  labs(title = "",y = "",
       x="")+
  facet_wrap(~End,scales = "free")+
  theme_bw()+
  scale_color_manual("",values=c("darkred","darkblue"))+
  theme(legend.position="top")
```